---
title: 'Simulated data for WCGNA demonstration'
author: "Samuel O Skinner"
date: "April 11, 2017"
output:
  html_document:
    number_sections: yes
---


```{r global_opts, echo=FALSE, cache=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4, fig.align='center',
                      echo=TRUE, warning=FALSE, message=FALSE,
                      cache=FALSE, autodep=TRUE)
# knitr::opts_knit$set(root.dir = "..")

## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 1, digits = 5)
```


# Summary


----- 

# R environment set-up

## Loading packages

```{r load_packages, cache=FALSE}
# Clean up the environment
rm(list = ls())
cleanup <- gc(verbose = FALSE)

# Load libraries I'll need here
library(MASS)
library(dplyr)
library(limma)
library(edgeR)
library(readr)
library(readxl)
library(ggplot2)
library(cowplot)
library(reshape2)
library(WGCNA)
library(gplots)
library(corrplot)
library(caret)
library(leaps)

library(glmnet)

# Packages for R markdown stuff
library(knitr)
library(shiny)

# Required for WGCNA to work properly
options(stringsAsFactors = FALSE)
```


``` {r helper_plotting_function, echo=FALSE}


plotMEvsTT <- function(module, feature) {

  modNames <- substring(names(MEs), 3)
  
  Trait_data <- datTraits[,colnames(datTraits)==feature]
  ME_counts <- MEs[, match(module, modNames)]
  
  out_plot <- ggplot(data.frame(x=ME_counts,
                                y=Trait_data),
                     aes(x=x,
                         y=y)) + 
    geom_point() +
    geom_smooth(method=lm) + 
    labs(x=paste(module, "module"),
         y=feature,
         main="Module vs. Trait")
  
  return(out_plot)

}


plotGSvsMM <- function(module, feature, MEs, datCounts) {

  modNames <- substring(names(MEs), 3)
  geneModuleMembership <- as.data.frame(cor(datCounts, MEs, use = "p"))
  MMPvalue <- as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples))
                    
  names(geneModuleMembership) = paste("MM", modNames, sep="");
  names(MMPvalue) = paste("p.MM", modNames, sep="");
  
  Trait_data <- datTraits[,colnames(datTraits)==feature]
  geneTraitSignificance = as.data.frame(cor(datCounts, Trait_data, use = "p"));
  GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));
  
  out_plot <- ggplot(data.frame(x=abs(geneModuleMembership[moduleColors==module, match(module, modNames)]),
                    y=abs(geneTraitSignificance[moduleColors==module, 1])),
         aes(x=x,
             y=y)) + 
    geom_point() +
    geom_smooth() + 
    labs(x=paste("MM in", module, "module"),
         y=paste("GS for ", feature),
         main="Module membership vs. gene significance")
  
  return(out_plot)

}



```




# Loading data

## Load count, annotation, metric, etc files.

```{r load_data}

# Read in file created during QC scripts
load("data/CAV-Cohort-2-RNAseq_Bcell_1_QC_20170209a.RData")

# This curDGE has already been filtered by Matt Dufort for poor or contaminated libraries and by me for low-count libraries
#    as well as transformed with voom
dim(t) # 28 obs. of 130 vars
dim(metriSum) # 378 obs of 5 vars
dim(curDGE)   # 11338 obs. of 28 vars


```

## Load in file with C-peptide decline and clinical parameters

```{r read in cpeptide decline file}

# Load in C-peptide decline data
cpepDat <- read.csv("../CAV-Cohort-2-DemographicData/data/CAV-Cohort-2_outcomes_parameters_20170221a.csv")

cpepSub <- cpepDat %>% select(c(participant_id,
                                study,
                                sex,
                                age_years,
                                bmi,
                                cpep_model_absdiff_1year,
                                cpep_model_absdiff_2year,
                                cpep_model_decayrate,
                                cpep_auc2hr_log_baseline))
cpepSub <- cpepSub[!duplicated(cpepSub$participant_id),]

annotDat <- annotDat %>% select(participant_id, libid)

cpepSub <- left_join(annotDat, cpepSub, by="participant_id")

```

As noticed by Matt during his QC scripts: There are no ABATE samples. 



# Visualize some of the clinical parameters

```{r plotting_the_c-peptide_baseline, fig.height=12, fig.width=12, echo=FALSE}

p1 <- ggplot(cpepSub, 
       aes(x=cpep_model_decayrate)) +
  geom_line(stat="density") +
  labs(x="Slope of log( fraction AUC )")

p2 <- ggplot(cpepSub, 
       aes(x=cpep_auc2hr_log_baseline,
           y=cpep_model_decayrate)) +
  geom_point() +
  labs(x="log(C-peptide at Baseline)",
       y="Slope of log( fraction AUC )")

p3 <- ggplot(cpepSub, 
       aes(x=sex,
           y=cpep_model_decayrate)) +
  geom_boxplot() +
  labs(x="Sex",
       y="Slope of log( fraction AUC )")

p4 <- ggplot(cpepSub, 
       aes(x=age_years,
           y=cpep_model_decayrate)) +
  geom_point() +
  labs(x="Age",
       y="Slope of log( fraction AUC )")

p5 <- ggplot(cpepSub, 
       aes(x=study,
           y=cpep_model_decayrate)) +
  geom_boxplot() +
  labs(x="Study",
       y="Slope of log( fraction AUC )")

p6 <- ggplot(cpepSub, 
       aes(x=bmi,
           y=cpep_model_decayrate)) +
  geom_point() +
  labs(x="BMI",
       y="Slope of log( fraction AUC )")


plot_grid(p1, p2, p3, p4, p5, p6, ncol=2)


```


# Analysis  
  

## Clustering on all cells

```{r visualize_clustering, fig.height=4, fig.width=10, echo=FALSE}



#Transpose for hclust
datCounts <- data.frame(v=t(curDGE$E))

sampleTree <- hclust(dist(datCounts), method = "average")

cpepSub$study_col <- ifelse(cpepSub$study=="START", "blue", "red")

# Plot the sample tree labeled by cell type
plotDendroAndColors(sampleTree,
                    colors = cpepSub$study_col,
                    main = "Sample dendrogram and trait heatmap")


```


1) Cell type  
Blue: START  
Red:  T1DAL  


## Perform WGCNA 

``` {r num_top_var_genes, echo=FALSE}

top_var_genes <- 7000

```


The first step is to select the most variable genes within cells. The number of genes does affect the result, and we need to determine a way to determine the optimal number. Currently, the number is set to `r top_var_genes`. 


### Select soft-threshold for cells

```{r choosing_thresholding_power, echo=FALSE}


curDGE_orig <- curDGE

#Select genes with the most variation 
geneVariances <- apply(curDGE$E,1,var)
curDGE <- curDGE[order(geneVariances, decreasing=TRUE),]
curDGE <- curDGE[1:top_var_genes,]


# Choose a set of soft-thresholding powers
powers <- c(c(1:10), seq(from = 12, to=20, by=2))

# Call the network topology analysis function
sft <- pickSoftThreshold(t(curDGE$E), powerVector = powers, verbose=0, networkType = "signed")
sft$x <- sft$fitIndices[,1]
sft$y <- -sign(sft$fitIndices[,3])*sft$fitIndices[,2]
sft$z <- sft$fitIndices[,5]

#Adjacency matrix
softPower = 12;
adjacency = adjacency(t(curDGE$E), power = softPower, type = "signed");

# Turn adjacency into topological overlap
TOM = TOMsimilarity(adjacency, verbose=0);
dissTOM = 1-TOM


# Scale-free topology fit index as a function of the soft-thresholding power
pl1 <- ggplot(data.frame(sft),
       aes(x=x,
           y=y,
           label=powers)) + 
  geom_text() + 
  geom_abline(slope=0, 
              intercept=sft$y[softPower],
              color="red") +
  labs(x="Soft Threshold (power)",
       y="Scale Free Topology Model Fit,signed R^2",
       main="Scale independence")

# Mean connectivity as a function of the soft-thresholding power
pl2 <- ggplot(data.frame(sft),
       aes(x=x,
           y=z,
           label=powers)) + 
  geom_text() + 
  geom_abline(slope=0, 
              intercept=sft$z[softPower],
              color="red") +
  labs(x="Soft Threshold (power)",
       y="Mean Connectivity",
       main="Mean connectivity")

plot_grid(pl1, pl2, ncol=2)

```


The soft threshold for this data subset is selected to be `r softPower`.  
  
  

### Perform clustering to create modules & merge similar modules

First, calculate the co-expression similarity, adjacency, topological overlap matrix (TOM), and dissimilarity.  Then perform heirarchical clustering on dissimilarity based on topological overlap.  
  
Second, to merge eigengenes, use the height of 0.25, corresponding to a correlation of 0.75.  
  
Third, filter out genes that have no module membership above 0.50. Move all that don't to a white module.  


```{r plot_gene_tree, echo=FALSE}

# We like large modules, so we set the minimum module size relatively high:
minModuleSize = 30;

# Call the hierarchical clustering function
geneTree = hclust(as.dist(dissTOM), method = "average")
dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM,
                            deepSplit = 3, pamRespectsDendro = TRUE,
                            minClusterSize = minModuleSize,
                            verbose = 0)

dynamicColors = labels2colors(dynamicMods)
# white is the 27th color. if white's used, replace with the next color
if (max(dynamicMods) >= 27) {
  dynamicColors[dynamicColors == "white"] <- standardColors(max(dynamicMods)+1)[max(dynamicMods)+1]
}
# dynamicColors[dynamicColors=="grey"] <- "white"


# Calculate eigengenes
MEList = moduleEigengenes(t(curDGE$E), colors = dynamicColors, excludeGrey=TRUE)
MEs = MEList$eigengenes


# Calculate dissimilarity of module eigengenes
MEDiss = 1-cor(MEs);
# Cluster module eigengenes
METree = hclust(as.dist(MEDiss), method = "average");

#Cut dendrogram to merge modules if desired
MEDissThres = 0.25


# Plot the result
plot(METree, main = "Clustering of module eigengenes",
     xlab = "", sub = "")
# Plot the cut line into the dendrogram
abline(h=MEDissThres, col = "red")


# Call an automatic merging function
merge = mergeCloseModules(t(curDGE$E), dynamicColors, cutHeight = MEDissThres, verbose = 0, getNewUnassdME = FALSE, useAbs = FALSE)
# The merged module colors
moduleColors = merge$colors;
# Eigengenes of the new merged modules:
MEs = merge$newMEs



#Calculate modular membership
datKME=signedKME(t(curDGE$E), MEs)
# datKME = as.data.frame(abs(cor(t(curDGE$E), MEs))) % For unsigned analysis

#Fix KME row names. WGCNA will replace dashes with periods which creates issues later
#String db does not recognize gene names with periods instead of dashes
rownames(datKME) <- colnames(t(curDGE$E))
#Because it doesn't depend on module size, KME is recommended for comparing modules
mainKME <- rep(NA, length(moduleColors))
for (i in 1:length(moduleColors)){
  mainKME[i] <- ifelse(moduleColors[i]=="grey", 0, datKME[i,paste0("kME",moduleColors[i])]) 
}
datKME$mainKME <- mainKME


##Assign anything with KME < 0.50 to the white module
greyGenes <- which(datKME$mainKME < 0.75)
highKMEColors <- moduleColors
highKMEColors[greyGenes] <- "white"


plotDendroAndColors(geneTree, cbind(dynamicColors, moduleColors, highKMEColors),
                   c("Dynamic Tree Cut", "Merged dynamic", "k_{ME} Filtered"),
                   dendroLabels = FALSE, hang = 0.03,
                   addGuide = TRUE, guideHang = 0.05)

moduleColors <- highKMEColors

```



```{r gene expression plot, echo=FALSE, fig.width=10, fig.height=4}

ClusterData <- data.frame(gene=colnames(t(curDGE$E)),
                          symbol=curDGE$genes$symbol,
                          mainKME=datKME$mainKME,
                          color=highKMEColors)

ClusterCount <- t(curDGE$E)[,order(ClusterData$mainKME, decreasing=TRUE)]
ClusterData  <- ClusterData[order(ClusterData$mainKME, decreasing=TRUE),]

ClusterCount <- ClusterCount[,order(ClusterData$color)]
ClusterData  <- ClusterData[order(ClusterData$color),]

# ClusterCount <- ClusterCount[,ClusterData$color != "slategrey"]
# ClusterData  <- ClusterData[ClusterData$color != "grey",]

ClusterCount_sc <- data.frame(apply(ClusterCount, 2, function(x){(x-mean(x))/sd(x)}))



```


### Display modules

``` {r plots inspired by McKinney, fig.height=12, fig.width=10, echo=FALSE}

ClusterData <- data.frame(gene=colnames(t(curDGE$E)),
                          symbol=curDGE$genes$symbol,
                          mainKME=datKME$mainKME,
                          color=highKMEColors)

ClusterCount <- t(curDGE$E)[,geneTree$order]
ClusterData  <- ClusterData[geneTree$order,]



# Prep the raw gene expression data
ClusterCount_sc <- data.frame(apply(ClusterCount, 2, function(x){(x-mean(x))/sd(x)}))
l_sc <- mean(apply(ClusterCount_sc, 2, function(x){sqrt(sum(x^2))}))

samTree <- hclust(as.dist(1-cor(t(ClusterCount_sc))), method="average")
    
ClusterCount_sc <- ClusterCount_sc[samTree$order,]

# heatmap.2(as.matrix(ClusterCount_sc),
#           ColSideColors=ClusterData$color,
#           Rowv=FALSE,
#           Colv=FALSE,
#           dendrogram="none",
#           key=FALSE,
#           keysize = 0.1,
#           trace="none")

ClusterCount_sc$sample <- rownames(ClusterCount_sc)
ClusterCount_sc_tall <- melt(ClusterCount_sc)
ClusterCount_sc_tall_disp <- merge(ClusterCount_sc_tall, ClusterData, by.x="variable", by.y="gene", all.x=TRUE)
ClusterCount_sc_tall_disp$sample <- 1:dim(ClusterCount_sc)[1]

# Prep the eigengenes
MEs_disp <- MEs
colnames(MEs_disp) <- lapply(colnames(MEs_disp), function(x) {gsub("ME","",x)})
MEs_disp$sample <- rownames(ClusterCount_sc)
MEs_disp <- MEs_disp[samTree$order,]
MEs_disp_tall <- melt(MEs_disp)
MEs_disp_tall$sample <- 1:dim(ClusterCount_sc)[1]
MEs_disp_tall$value <- l_sc*MEs_disp_tall$value
MEs_disp_tall$color <- MEs_disp_tall$variable



ggplot(ClusterCount_sc_tall_disp,
       aes(x=sample,
           y=value)) +
  geom_line(aes(color = variable),
            alpha=0.25,
            size=0.3) +
  labs(y="Gene expression",
       x="Individual samples") + 
  theme(legend.position="none") + 
  facet_wrap(~color, ncol=2) +
  geom_line(data=MEs_disp_tall,
            aes(x=sample,
                y=value),
            color="red",
            size=1.5)
  
```


```{r correlation between the eigengenes, echo=FALSE}

corrplot(cor(MEs))


```




### Display top 200 genes in each module
Ordered by the correlation of gene to eigengene.

```{r display genes in each module, echo=FALSE}

ClusterData_disp <- ClusterData

ClusterData_disp <- ClusterData_disp[ClusterData_disp$color != "white",]
ClusterData_disp <- ClusterData_disp[order(ClusterData_disp$mainKME, decreasing = TRUE),]
ClusterData_disp <- ClusterData_disp[order(ClusterData_disp$color),]


A <- lapply(unique(ClusterData_disp$color), function(x) {print(x); temp<-ClusterData_disp[ClusterData_disp$color==x,]; print(head(temp$symbol,200))})

```



## Modeling c-peptide decline with features

Feature properties  

```{r feature histograms, fig.width=10, fig.height=8, echo=FALSE}
model_df <- cbind(cpepSub, MEs)

ggplot(melt(model_df),
       aes(x=value,
           color=study)) + 
  geom_line(stat="density") + 
  facet_wrap(~variable, scales="free")


```


### Model each eigengene individually

Model 1: cpep_model_decayrate ~ eigengene + bmi + sex + study + age_years + cpep_auc2hr_log_baseline  
Model 2: cpep_model_decayrate ~ eigengene + study + age_years + cpep_auc2hr_log_baseline 
```{r individual eigengene models, echo=FALSE}

model_df <- cbind(cpepSub, MEs)

color = colnames(MEs)[1]

fit1_p_value <- list()
fit1_MEp_value <- list()
fit2_p_value <- list()
fit2_MEp_value <- list()

for (color in colnames(MEs)) {
  
  f1 <- as.formula(paste("cpep_model_decayrate ~ ", color, " + bmi + sex + study + age_years + cpep_auc2hr_log_baseline"))
  fit1 <- lm(f1, data=model_df)
  fit1_f <- summary(fit1)$fstatistic
  fit1_p_value[color] <- ifelse(is.null(fit1_f), NA, pf(fit1_f[1],fit1_f[2],fit1_f[3],lower.tail=FALSE))
  fit1_MEp_value[color] <- summary(fit1)$coefficients[2,4]
  fit1_coef <- coef(summary(fit1))[,4]
  
  f2 <- as.formula(paste("cpep_model_decayrate ~ ", color, " + study + age_years + cpep_auc2hr_log_baseline"))
  fit2 <- lm(f2, data=model_df)
  fit2_f <- summary(fit2)$fstatistic
  fit2_p_value[color] <- ifelse(is.null(fit2_f), NA, pf(fit2_f[1],fit2_f[2],fit2_f[3],lower.tail=FALSE))
  fit2_MEp_value[color] <- summary(fit1)$coefficients[2,4]
  fit2_coef <- coef(summary(fit2))[,4]
  
}

kable(data.frame(cbind(fit1_p_value,fit1_MEp_value, fit2_p_value, fit2_MEp_value)))

```


### Various regression methods to model cpeptide decline with eigengenes and demo parameters

This modeling is first performed using all demographic parameters:  
Model: cpep_model_decayrate ~ Eigengenes + bmi + sex + study + age_years + cpep_auc2hr_log_baseline  

#### Linear regression with all eigengenes

```{r model cpep decline with eigengenes and demographic paramters, echo=FALSE}

model_df <- cbind(cpepSub, MEs)

color_str <- paste(colnames(MEs), collapse=" + ")

f <- as.formula(paste("cpep_model_decayrate ~ ", color_str, " + bmi + sex + study + age_years + cpep_auc2hr_log_baseline"))
fit <- lm(f, data=model_df)
fit_f <- summary(fit)$fstatistic
fit_p_value <- ifelse(is.null(fit_f), NA, pf(fit_f[1],fit_f[2],fit_f[3],lower.tail=FALSE))
fit_coef <- coef(summary(fit))[,4]

summary(fit)

```

#### Backward step-wise regression

```{r back step, echo=FALSE}

step_fit <- step(fit, trace=FALSE)
step_f <- summary(step_fit)$fstatistic
step_p_value <- ifelse(is.null(step_f), NA, pf(step_f[1],step_f[2],step_f[3],lower.tail=FALSE))
step_coef <- coef(summary(step_fit))[,4]

summary(step_fit)

# plot(step_fit)

```


#### Forward step-wise regression
Starting from: cpep_model_decayrate ~ study + age_years + cpep_auc2hr_log_baseline

```{r forward stepwise, echo=FALSE}

f <- as.formula("cpep_model_decayrate ~ study + age_years + cpep_auc2hr_log_baseline")
forw_fit <- step(lm(f, data=model_df), 
                 trace=FALSE, 
                 direction="forward", 
                 scope=list(lower=as.formula("cpep_model_decayrate ~ study + age_years + cpep_auc2hr_log_baseline"),
                            upper=as.formula("cpep_model_decayrate ~ .")))
forw_f <- summary(forw_fit)$fstatistic
forw_p_value <- ifelse(is.null(forw_f), NA, pf(forw_f[1],forw_f[2],forw_f[3],lower.tail=FALSE))
forw_coef <- coef(summary(forw_fit))[,4]

summary(forw_fit)

```

#### All subset regression

```{r all subset regression, echo=FALSE}

f <- as.formula(paste("cpep_model_decayrate ~ ", color_str, " + bmi + sex + study + age_years + cpep_auc2hr_log_baseline"))
reg1 <- regsubsets(f, 
                   data = model_df, 
                   nvmax = 15)

plot(reg1)

```



#### Elastic net regression

```{r en model cpep decline with eigengenes and demographic paramters, fig.width=10, fig.height=5, echo=FALSE}


cpepSub2 <- cpepSub %>% select(sex, study, age_years, bmi, cpep_auc2hr_log_baseline)
cpepSub2$sex <- ifelse(cpepSub2$sex=="M", 0, 1)
cpepSub2$study <- ifelse(cpepSub2$study=="START", 0, 1)

x <- as.matrix(cbind(cpepSub2, MEs))
y <- as.matrix(cpepSub$cpep_model_decayrate)

# Use glmnet
fit1=cv.glmnet(x, y, standardize=TRUE, family="gaussian", nfolds=nrow(x))
plot(fit1)

```

Check which features are included with lambda value.
```{r coeff glmnet}

coef(fit1, s = exp(-8.5))

```
Scan lambda and alpha.
```{r using caret, fig.width=10, fig.height=6, echo=FALSE}

# Use caret
set.seed(104) # 104 is a great seed
cctrl1 <- trainControl(method = "cv", number = 10)
test_class_cv_model <- train(x, as.numeric(y),
                             method = "glmnet",
                             trControl = cctrl1,
                             metric = "RMSE",
                             preProc = c("center", "scale"),
                             tuneGrid = expand.grid(.alpha = seq(0, 1,.1),
                                                    .lambda = exp(seq(-15,-5,.25))))


ggplot(test_class_cv_model$results,
       aes(x = lambda,
           y = RMSE,
           color = as.factor(alpha))) + 
  geom_point() + 
  geom_line() + 
  scale_x_log10() + 
  labs(x = "Regularization Parameter (lambda)",
       y = "RMSE (Cross-Validation)")


# Use caret
set.seed(104) # 104 is a great seed
cctrl1 <- trainControl(method = "cv", number = 10)
test_class_cv_model <- train(x, as.numeric(y),
                             method = "glmnet",
                             trControl = cctrl1,
                             metric = "RMSE",
                             preProc = c("center", "scale"),
                             tuneGrid = expand.grid(.alpha = seq(.9, 1,.1),
                                                    .lambda = exp(seq(-11,-9.5,.25))))


ggplot(test_class_cv_model$results,
       aes(x = lambda,
           y = RMSE,
           color = as.factor(alpha))) + 
  geom_point() + 
  geom_line() + 
  scale_x_log10() + 
  labs(x = "Regularization Parameter (lambda)",
       y = "RMSE (Cross-Validation)")

plot(varImp(test_class_cv_model,scale=F))

```


#### Recursive feature elimination

```{r recursuve feature elimination with caret, echo=FALSE}


cpepSub2 <- cpepSub %>% select(sex, study, age_years, bmi, cpep_auc2hr_log_baseline)
cpepSub2$sex <- ifelse(cpepSub2$sex=="M", 0, 1)
cpepSub2$study <- ifelse(cpepSub2$study=="START", 0, 1)

x <- as.matrix(cbind(cpepSub2, MEs))
# y <- as.matrix(cpepSub$cpep_model_absdiff_2year)
# y <- as.matrix(-365.24*cpepSub$cpep_model_decayrate)
y <- as.matrix(cpepSub$cpep_model_decayrate)


cor_features <- findCorrelation(cor(x))
print(colnames(x)[cor_features])

if (length(cor_features)<1) {dat_x <- x} else {dat_x <- x[,-cor_features]}
# dat_x <- x

set.seed(48)

subsets <- seq(2,100,1)

ctrl <- rfeControl(functions = lmFuncs,
                   method = "repeatedcv",
                   repeats = 5,
                   rerank = TRUE,
                   verbose = FALSE)

lmProfile <- rfe(dat_x, 
                 y,
                 sizes = subsets,
                 rfeControl = ctrl)

print(lmProfile)

# plot the results
plot(lmProfile, type=c("g", "o"))



```


### Simple lm model from the consensus of different regression strategies


```{r consensus model}


model_df <- cbind(cpepSub, MEs)

color_str <- paste(colnames(MEs), collapse=" + ")

f <- as.formula(paste("cpep_model_decayrate ~ MElightcyan + MEmagenta + bmi + sex + study + age_years + cpep_auc2hr_log_baseline"))
fit <- lm(f, data=model_df)
fit_f <- summary(fit)$fstatistic
fit_p_value <- ifelse(is.null(fit_f), NA, pf(fit_f[1],fit_f[2],fit_f[3],lower.tail=FALSE))
fit_coef <- coef(summary(fit))[,4]

summary(fit)



ggplot(data.frame(x=predict(test_class_cv_model, x),
                  y=y),
       aes(x=x,y=y)) +
  geom_point() +
  geom_abline(slope=1, intercept=0, color="red")

```




## Associations of features

### Association of eigengenes with categorical parameters


```{r associating the categorical parameters with eigengenes, echo=FALSE}

datTraits <- cpepSub %>% select(c(sex, study))
wt.dat <- cbind(MEs, datTraits)

wt_sex <- list()
wt_stu <- list()

for (color in colnames(MEs)) {
  
  f <- as.formula(paste(color," ~ sex", sep=""))
  wt_sex[color] <- wilcox.test(f, data=wt.dat)$p.value
    
  f <- as.formula(paste(color," ~ study", sep=""))
  wt_stu[color] <- wilcox.test(f, data=wt.dat)$p.value
  
}

wt_sex <- unlist(wt_sex)
wt_stu <- unlist(wt_stu)


```
P-values of a Mann-Whitney test between eigengene ~ sex.  
`r kable(data.frame(pvalue=wt_sex, adj.pvalue=p.adjust(wt_sex, method="BH")))`

P-values of a Mann-Whitney test between eigengene ~ study.  
`r kable(data.frame(pvalue=wt_stu, adj.pvalue=p.adjust(wt_stu, method="BH")))` 

### Correlation of eigengenes with continuous demographic parameters

Correlations
```{r correlate_eigengenes, fig.width=10, fig.height=10, echo=FALSE}

datTraits <- cpepSub %>% select(c(age_years, 
                                  bmi,
                                  cpep_model_decayrate, 
                                  cpep_model_absdiff_1year,
                                  cpep_model_absdiff_2year,
                                  cpep_auc2hr_log_baseline))


# Define numbers of genes and samples
nSamples = nrow(t(curDGE$E));

# Recalculate MEs with color labels
moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);



# Correlation plot
corrplot(moduleTraitCor, p.mat = moduleTraitPvalue, insig="p-value", sig.level=0)

```


Displaying the significant correlations (p < 0.15) as scatter-plots.

```{r gene_significance_and_module_membership, fig.width=15, fig.height=8, echo=FALSE}

# Now, filter by for significant correlation sets
sig.df <- melt(moduleTraitPvalue)
sig.df <- sig.df[sig.df$value < 0.15,]

datTraits <- cbind(datTraits, MEs)


plt <- list()
for (i in 1:dim(sig.df)[1]) {
  
  feature  <- as.character(sig.df[i,2])
  module   <- strsplit(as.character(sig.df[i,1]),"ME")[[1]][2]
  plt[[i]] <- plotMEvsTT(module=module, feature=feature)
  
}

plot_grid(plotlist=plt)

```



## Enrichment analysis  

### Immune-pathway gene set enrichment analysis for modules of interest

```{r GSEA, echo=FALSE}


labelR <- curDGE_orig$genes
labelR <- labelR[!duplicated(labelR$symbol),]
labelR <- left_join(labelR, data.frame(symbol=curDGE$genes$symbol, color=moduleColors), by="symbol")
labelR$color[is.na(labelR$color)] <- "white"

GSEA <- userListEnrichment(labelR$symbol, 
                           labelR$color, 
                           fnIn=NULL,
                           omitCategories = "white",
                           useImmunePathwayLists = TRUE,
                           outputCorrectedPvalues=TRUE)

GSEA_sig <- paste(GSEA$pValues[GSEA$pValues$CorrectedPvalues<0.15,]$InputCategories, 
                  GSEA$pValues[GSEA$pValues$CorrectedPvalues<0.15,]$UserDefinedCategories, 
                  sep=" -- ")


kable(GSEA$pValues[GSEA$pValues$CorrectedPvalues<0.15,])

GSEA$ovGenes[GSEA_sig]


```


### Gene ontology analysis for modules of interest

```{r GO_analysis, echo=FALSE}

GOenr = GOenrichmentAnalysis(moduleColors, curDGE$genes$entrez, organism = "human", nBestP = 10, verbose=0)

tab = GOenr$bestPTerms[[4]]$enrichment


keepCols = c(1, 2, 5, 6, 7, 12, 13)
screenTab = tab[, keepCols];
# Round the numeric columns to 2 decimal places:
numCols = c(3, 4);
screenTab[, numCols] = signif(apply(screenTab[, numCols], 2, as.numeric), 2)
# Truncate the the term name to at most 40 characters
screenTab[, 7] = substring(screenTab[, 7], 1, 40)
# Shorten the column names:
colnames(screenTab) = c("module", "size", "p-val", "Bonf", "nInTerm", "ont", "term name");
rownames(screenTab) = NULL;
# Set the width of R’s output. The reader should play with this number to obtain satisfactory output.
options(width=95)
# Finally, display the enrichment table:
screenTab

```




## Save eigengenes for use in module-based model
```{r save eigenganes, echo=FALSE}

rownames(MEs) <- cpepSub$participant_id

modules <- list()
for (module in colnames(MEs)) {

  modules[[module]] <- ClusterData[ClusterData$color == strsplit(module,"ME")[[1]][2],]$symbol

}

save(list=c("MEs", "modules"), file="data/CAV-Cohort-2-RNAseq_Bcell_WGCNAModules_20170405a.RData")


# Save a csv for Cate to check out
write.csv(data.frame(moduleID = names(unlist(modules)),
                     gene_symbol = unlist(modules)),
          quote=FALSE,
          row.names=FALSE,
          file="data/CAV-Cohort-2-RNAseq_Bcell_WGCNAModules_20170405a.csv")

```

